Leveraging Diversity to Find Bugs in JavaScript Engines

### 2019-02-21 paper/paper.tex
related work + conclusions + spell checking

%% It automates test generation in scenarios where multiple
%% implementations of a system exist. DT leverages diversity across
%% system's implementations to detect anomalous behavior.
\emph{Related Ideas.}~The idea of test set diversity dates back to the
eighties~\cite{white-cohen-tse1980, ostrand-balcer-1988}. In constrast
to prior work on test set diversity, this paper explores diversity of
sources of test cases and diversity of implementations with the goal
of finding bugs. Section~\ref{sec:related-work} elaborates related
work on this topic. Differential Testing~\cite{Brumley-etal-ss07} (DT)
has been applied in a variety of contexts to find
bugs~\cite{Yang-etal-pldi11,Chen-etal-fse2015,Argyros-etla-ccs16,Chen-etal-pldi16,petsios-etal-sp2017,SivakornAPKJ17,Zhang:2017:ATD:3097368.3097448}.
It has shown to be specially practical in scenarios where the
observation of difference gives a strong signal of a real problem. For
example, Mozilla runs JS files against different configurations of
a given build of their \smonkey\ engine (\eg{}, trying to enable or
not eager JIT compilation\footnote{These files are created with
  the grammar-based fuzzer jsfunfuzz~\cite{jsfunfuzz}. Look for option
  ``compare\_jit'' from funfuzz.}). A positive aspect of the approach is that
it can be fully automated---as only one engine is used, the outcomes
of the test in both configurations are expected to be identical. The
Mozilla team uses this approach since 2002; they have been able to
find over 270 bugs since then~\cite{jsfunfuzz-at-mozilla}, including
security bugs. Cross-engine differential testing, in contrast, has not
been widely popularized. One possible reason is that it is still not
practical to fully automate the technique. In contrast to other
applications of differential testing, a number of legitimate reasons
exist, other than a bug, for a test execution to manifest discrepancy
(see Tables~\ref{fig:piecharts-transplantation} and
~\ref{tab:false-positives}). To sum, variations of these ideas have
been explored before in different contexts.  The goal of the paper is
to assess the ability of the techniques aforementioned in finding
bugs on \javascript\ engines.

@@

For the classification of ``code'' vs. ``not code'', we used a
pipeline of two Neural Networks (NN), a popular design for solving NLP
classification problems~\cite{kusner2015word}. The first net in the
pipeline takes as input an arbitrary sentence and produces on output a
characterization vector for that sentence. The second net in the
pipeline takes that vector on input and produces a yes/no answer,
determining whether or not the input sentence was code.  During the
training phase, the second neural net in the pipeline takes
additionaly on input the value for the class attribute for the sample;
in this case, ``code'' or ``not code''. More in detail, for the first
net, we used word2vec~\cite{mikolov2013distributed}, a popular NLP
technique to produce word embeddings. A word embedding is a mapping of
words to vectors of real numbers. For the second net, we used a
multi-layer perceptron~\cite{Rumelhart:1986:LIR:104279.104293} to
infer the probability of the input belonging or not to the class. The
classifier labels the input input as code if the predicted probability
of the input being code is 0.7 or higher. We used a corpus with 25K
samples of English paragraphs and 25K snippets of \js{} code to train
and test the classifier and obtained an accuracy of 98\%. This
classifier is publicly available from our website as a separate
component.

@@

\begin{wraptable}[12]{r}{0.23\textwidth}
  \vspace{-1ex}
  \small
  \centering
  \caption{\label{fig:falsepositives}\label{fig:truepositives}\label{fig:piecharts-transplantation}Distribution
    of False (FP) and True Positives (TP).}
  \setlength{\tabcolsep}{3pt}      
  \renewcommand*{\arraystretch}{0.9}
  \begin{tabular}{ccr}
    \toprule
    & source &  \#\\
    \midrule
    \multirow{5}{*}{FP} & Undefined Behavior & \noTransUndefined{} \\
    & Timeout/OME & \noTransTimeout{} \\
    & Not Implemented & \noTransNotImplemented{} \\
    & Non-Standard Element & \noTransNonStandard{} \\
    & Other & \noTransOther{} \\
    \midrule
    \multirow{2}{*}{TP} & Duplicate & \noTransTPDuplicated{} \\
    & Bug & \noTransTPBugs{} \\
    \bottomrule
  \end{tabular}
\end{wraptable}

\Igor{
  Os numeros da Tabela 5 correspondem ao numero de arquivos que falharam 
  em test transplantation (nesse caso \failuresTestTransDistictFiles{}).
  houveram \failuresTestTrans{} falhas que correspondem aos dados da Tabela 4.
}

Table~\ref{fig:piecharts-transplantation} shows the distribution of
False Positives (FPs) and True Positives (TPs). Considering false
positives, ``Undefined Behavior'' was the most predominant
source. Considering true positives, we found a reasonable number of
duplicate reports, but not high enough to justify attempting to
automate the detection of duplicates.

@@

The object property \CodeIn{valueOf} stores a function that returns a
primitive value identifying the target object~\cite{valueof}. The
original version of this code returns an empty string whereas the
version of the code modified by the \radamsa{} fuzzer~\cite{radamsa}
returns a string representation of a null character (\CodeIn{NUL})\Comment{ in
ascii}. The unary plus expression ``\CodeIn{+a}", used in the
assertion, is equivalent to the operation
\CodeIn{ToNumber(a.valueOf())} that converts a string to a number,
otherwise the operation returns NaN (Not a
Number)~\cite{unary-plus}. This test fails in all engines but
\chakra{}. For all three engines the string cannot be parsed as an
hexadecimal. As such, they produce a NaN as result and the test fails
as expected. \chakra{}, in contrast, incorrectly converts the string
to zero, making the test to pass. As Table~\ref{tab:bugs} shows, the
\chakra{} team fixed the issue soon after we reported the problem.

@@

The main lessons we learned from this study are as follows: 1)~Even
for software projects with fairly clear specifications, as the case of
\javascript{}~\cite{ecmas262-spec}, there is likely to be (a lot of)
variation between different implementations and, therefore,
opportunities for bugs. 2)~Not only multiple different implementations
can be leveraged in differential testing, but differences in test
suites can also be important. 3)~Finding functional/non-crash bugs
with differential testing is feasible on real, complex, widely used
pieces of software. 4)~Further reducing cost of inspection is an
important problem. Although the inspection activity was not
uninterrupted, it is safe to say that each warning required a
substantial amount of time to analyze for potential false alarms. In
fact, many \hi\ warnings reported with differential testing were not
analyzed. In determining cost, we observed, from experience, that the
complexity of the \js\ specifications that the original test covers
increases cost of diagnosing (as developers need to read and
understand those) and the availability of alternative implementations
(for the cases warnings are revealed through differential testing)
reduces the cost of diagnosis. We prefer to see such problem as an
opportunity for future research. For example, applying learning
techniques to prioritize the warnings more likely to be faulty may be
a promising avenue to explore. Recall that most of those warnings are
false positives.  5)~We learned that reporting real bugs is a great
way to train (and encourage) students in software testing. Students
praised the experience of diagnosing failures, understanding part of
the specs (as needed), writing bug reports, participating in
discussions on issue trackers, and observing the change of
status. That was a relatively self-contained hands-on activity that
enabled students to engage in a real-life serious industrial project.

@@

\subsection{Differential Testing}
Several different applications of differential testing have been
proposed in recent years. Chen and
colleagues~\cite{Chen:2018:RDT:3180155.3180226} recently proposed a
technique to generate X.509 certificates based on Request For
Proposals (RFC) as specification with the goal of detecting bugs in
different SSL/TLS implementations. Those bugs can comprimise security
of servers which rely on these certificates to properly authenticate
the parties involved in a communication session. Lidbury and
colleagues~\cite{Lidbury:2015:MCF:2737924.2737986} and Donaldson and
colleagues~\cite{Donaldson:2017:ATG:3152284.3133917} have been
focusing on finding bugs in programs for graphic cards (\eg{},
OpenCL). These programs use the Single-Instruction Multiple-Data
(SIMD) programming abstraction and typically run on GPUs.  Perhaps the
application of differential testing that received most attention to
date was compiler testing. In 1972, Purdom~\cite{Purdom1972} proposed
the use of a generator of sentences from grammars to test correctness
of automatically generated parsers. After that, significant progress
has been made. Lammel and Shulte proposed Geno to cross-check XPath
implementations using grammar-based testing with controllable
combinatorial coverage~\cite{10.1007/11754008_2}. Yang and
colleagues~\cite{Yang:2011:FUB:1993498.1993532} proposed CSmith to
randomly create C programs from a grammar, for a subset of C, and then
check the output of these programs in different compilers (\eg{}, GCC
and LLVM). Le and colleagues~\cite{Le:2014:CVV:2594291.2594334}
proposed ``equivalence modulo inputs'', which creates variants of
program which should have equivalent behavior compared to the
original, but for which the compiler manifests
discrepancy. Differential testing has also been applied to test
refactoring engines~\cite{Daniel:2007:ATR:1287624.1287651}, to test
symbolic engine implementations~\cite{Kapus:2017:ATS:3155562.3155636},
and to test disassemblers and binary
lifters~\cite{Paleari:2010:NDD:1831708.1831741,Kim:2017:TIR:3155562.3155609}. All
in all, it has shown to be flexible and effective for a wide range of
applications. Surprisingly, not much work has been done on
differential testing of \js\ engines. Mozilla uses differential
testing to look for discrepanices across different configurations of
the same version of its \smonkey\ engine (using the ``compare\_jit''
flag of jsfunfuzz~\cite{jsfunfuzz}) whereas we focus on discrepancy
across engines. Patra and Pradel evaluated their language-agnostic
fuzzing strategy using differential testing. Their focuses on finding
differential bugs across multiple
browsers~\cite{patra2016learning}. As such they specialized their
fuzzer to HTML and JS (see Section~\ref{sec:testing-js-engines}). In
contrast to Patra and Pradel, we did not propose new techniques; our
contribution was empirical. \Fix{differential testing of debuggers...}

@@

\Fix{
although more research needs be done to automate
parts of the inspection process (\eg{}, reduction of false alarms),}

JavaScript is a very popular language. For that reason, bugs in engine
implementations often affect lots of people and organizations.
Implementing correct engines is challenging because the specification
is intentionally incomplete and evolves frequently. Finding bugs is
challenging for similar reasons. We conducted a study to evaluate two
simple diversity-aware techniques--test transplantation and
cross-engine differential testing. The first technique runs test
suites of a given engine in another engine. The second technique
fuzzes existing inputs and then compares the output produced by
different engines with a differential oracle.

Overall, we reported \totalBugsReported\ bugs in this study. Of which,
\totalBugsConfirmed\ were confirmed by developers and
\totalBugsFixed\ were fixed. Although more work is necessary to
reduce cost of manual analysis, we found that both techniques we used
are effective at finding bugs in
\js\ engines. Section~\ref{sec:lessons} elaborates on the lessons
learned from this study. The result data, infrastructure, and
experimental scripts we prepared to run this study are available upon
request. Future work includes mining test inputs from alternative
sources (\eg{}, bug trackers) and predicting likelihood of false
positives as to guide human effort on the most promising warnings
first.

@@

%%  LocalWords:  formedness Embeddings NLP Convolutional dataset

>>>>

%% It automates test generation in scenarios where multiple
%% implementations of a system exist. DT leverages diversity across
%% system's implementations to detect anomalous behavior.
\emph{Related Ideas.}~The idea of test set diversity dates back to the
eighties~\cite{white-cohen-tse1980, ostrand-balcer-1988}. In contrast
to prior work on test set diversity, this paper explores diversity of
sources of test cases and diversity of implementations with the goal
of finding bugs. Section~\ref{sec:related-work} elaborates related
work on this topic. Differential Testing~\cite{Brumley-etal-ss07} (DT)
has been applied in a variety of contexts to find
bugs~\cite{Yang-etal-pldi11,Chen-etal-fse2015,Argyros-etla-ccs16,Chen-etal-pldi16,petsios-etal-sp2017,SivakornAPKJ17,Zhang:2017:ATD:3097368.3097448}.
It has shown to be specially practical in scenarios where the
observation of difference gives a strong signal of a real problem. For
example, Mozilla runs JS files against different configurations of
a given build of their \smonkey\ engine (\eg{}, trying to enable or
not eager JIT compilation\footnote{These files are created with
  the grammar-based fuzzer jsfunfuzz~\cite{jsfunfuzz}. Look for option
  ``compare\_jit'' from funfuzz.}). A positive aspect of the approach is that
it can be fully automated---as only one engine is used, the outcomes
of the test in both configurations are expected to be identical. The
Mozilla team uses this approach since 2002; they have been able to
find over 270 bugs since then~\cite{jsfunfuzz-at-mozilla}, including
security bugs. Cross-engine differential testing, in contrast, has not
been widely popularized. One possible reason is that it is still not
practical to fully automate the technique. In contrast to other
applications of differential testing, a number of legitimate reasons
exist, other than a bug, for a test execution to manifest discrepancy
(see Tables~\ref{fig:piecharts-transplantation} and
~\ref{tab:false-positives}). To sum, variations of these ideas have
been explored before in different contexts.  The goal of the paper is
to assess the ability of the techniques aforementioned in finding
bugs on \javascript\ engines.

@@

For the classification of ``code'' vs. ``not code'', we used a
pipeline of two Neural Networks (NN), a popular design for solving NLP
classification problems~\cite{kusner2015word}. The first net in the
pipeline takes as input an arbitrary sentence and produces on output a
characterization vector for that sentence. The second net in the
pipeline takes that vector on input and produces a yes/no answer,
determining whether or not the input sentence was code.  During the
training phase, the second neural net in the pipeline takes
additionally on input the value for the class attribute for the sample;
in this case, ``code'' or ``not code''. More in detail, for the first
net, we used word2vec~\cite{mikolov2013distributed}, a popular NLP
technique to produce word embeddings. A word embedding is a mapping of
words to vectors of real numbers. For the second net, we used a
multi-layer perceptron~\cite{Rumelhart:1986:LIR:104279.104293} to
infer the probability of the input belonging or not to the class. The
classifier labels the input input as code if the predicted probability
of the input being code is 0.7 or higher. We used a corpus with 25K
samples of English paragraphs and 25K snippets of \js{} code to train
and test the classifier and obtained an accuracy of 98\%. This
classifier is publicly available from our website as a separate
component.

@@

\begin{wraptable}[12]{r}{0.23\textwidth}
  \vspace{-3ex}
  \small
  \centering
  \caption{\label{fig:falsepositives}\label{fig:truepositives}\label{fig:piecharts-transplantation}Distribution
    of False (FP) and True Positives (TP).}
  \setlength{\tabcolsep}{3pt}      
  \renewcommand*{\arraystretch}{0.9}
  \begin{tabular}{ccr}
    \toprule
    & source &  \#\\
    \midrule
    \multirow{5}{*}{FP} & Undefined Behavior & \noTransUndefined{} \\
    & Timeout/OME & \noTransTimeout{} \\
    & Not Implemented & \noTransNotImplemented{} \\
    & Non-Standard Element & \noTransNonStandard{} \\
    & Other & \noTransOther{} \\
    \midrule
    \multirow{2}{*}{TP} & Duplicate & \noTransTPDuplicated{} \\
    & Bug & \noTransTPBugs{} \\
    \bottomrule
  \end{tabular}
\end{wraptable}

%% \Igor{
%%   Os numeros da Tabela 5 correspondem ao numero de arquivos que falharam 
%%   em test transplantation (nesse caso \failuresTestTransDistictFiles{}).
%%   houveram \failuresTestTrans{} falhas que correspondem aos dados da Tabela 4.
%% }

Table~\ref{fig:piecharts-transplantation} shows the distribution of
False Positives (FPs) and True Positives (TPs). The sum of the numbers
in this table correspond to the number of files that manifested
failures, \ie{}, \failuresTestTransDistictFiles{}. Considering false
positives, ``Undefined Behavior'' was the most predominant
source. Considering true positives, we found a reasonable number of
duplicate reports, but not high enough to justify attempting to
automate the detection of duplicates.

@@

The object property \CodeIn{valueOf} stores a function that returns a
primitive value identifying the target object~\cite{valueof}. The
original version of this code returns an empty string whereas the
version of the code modified by the \radamsa{} fuzzer~\cite{radamsa}
returns a string representation of a null character (\CodeIn{NUL})\Comment{ in
ascii}. The unary plus expression ``\CodeIn{+a}", used in the
assertion, is equivalent to the operation
\CodeIn{To Number(a.valueOf())} that converts a string to a number,
otherwise the operation returns NaN (Not a
Number)~\cite{unary-plus}. This test fails in all engines but
\chakra{}. For all three engines the string cannot be parsed as an
hexadecimal. As such, they produce a NaN as result and the test fails
as expected. \chakra{}, in contrast, incorrectly converts the string
to zero, making the test to pass. As Table~\ref{tab:bugs} shows, the
\chakra{} team fixed the issue soon after we reported the problem.

@@

The main lessons we learned from this study are as follows: 1)~Even
for software projects with fairly clear specifications, as the case of
\javascript{}~\cite{ecmas262-spec}, there is likely to be (a lot of)
variation between different implementations and, therefore,
opportunities for bugs. 2)~Not only multiple different implementations
can be leveraged in differential testing, but differences in test
suites can also be important. 3)~Finding functional/non-crash bugs
with differential testing is feasible on real, complex, widely used
pieces of software. 4)~Further reducing cost of inspection is an
important problem. Although the inspection activity was not
uninterrupted, it is safe to say that each warning required a
substantial amount of time to analyze for potential false alarms. In
fact, many \hi\ warnings reported with differential testing were not
analyzed. In determining cost, we observed, from experience, that the
complexity of the \js\ specifications that the original test covers
increases cost of diagnosing (as developers need to read and
understand those) and the availability of alternative implementations
(for the cases warnings are revealed through differential testing)
reduces the cost of diagnosis. We prefer to see such problem as an
opportunity for future research. For example, applying learning
techniques to prioritize the warnings more likely to be faulty (in the
spirit of the work of Chen and
colleagues~\cite{Chen:2017:LPT:3097368.3097451}) may be a promising
avenue to explore. Recall that the rate of true positives of the
techniques we studied is rather small.  5)~We learned that reporting
real bugs is a great way to train (and encourage) students in software
testing. Students praised the experience of diagnosing failures,
understanding part of the specs (as needed), writing bug reports,
participating in discussions on issue trackers, and observing the
change of status. That was a relatively self-contained hands-on
activity that enabled students to engage in a real-life serious
industrial project.

@@

\subsection{Differential Testing}
Several different applications of differential testing have been
proposed in recent years. Chen and
colleagues~\cite{Chen:2018:RDT:3180155.3180226} recently proposed a
technique to generate X.509 certificates based on Request For
Proposals (RFC) as specification with the goal of detecting bugs in
different SSL/TLS implementations. Those bugs can compromise security
of servers which rely on these certificates to properly authenticate
the parties involved in a communication session. Lidbury and
colleagues~\cite{Lidbury:2015:MCF:2737924.2737986} and Donaldson and
colleagues~\cite{Donaldson:2017:ATG:3152284.3133917} have been
focusing on finding bugs in programs for graphic cards (\eg{},
OpenCL). These programs use the Single-Instruction Multiple-Data
(SIMD) programming abstraction and typically run on GPUs.  Perhaps the
application of differential testing that received most attention to
date was compiler testing. In 1972, Purdom~\cite{Purdom1972} proposed
the use of a generator of sentences from grammars to test correctness
of automatically generated parsers. After that, significant progress
has been made. Lammel and Shulte proposed Geno to cross-check XPath
implementations using grammar-based testing with controllable
combinatorial coverage~\cite{10.1007/11754008_2}. Yang and
colleagues~\cite{Yang:2011:FUB:1993498.1993532} proposed CSmith to
randomly create C programs from a grammar, for a subset of C, and then
check the output of these programs in different compilers (\eg{}, GCC
and LLVM). Le and colleagues~\cite{Le:2014:CVV:2594291.2594334}
proposed ``equivalence modulo inputs'', which creates variants of
program which should have equivalent behavior compared to the
original, but for which the compiler manifests
discrepancy. Differential testing has also been applied to test
refactoring engines~\cite{Daniel:2007:ATR:1287624.1287651}, to test
symbolic engine implementations~\cite{Kapus:2017:ATS:3155562.3155636},
to test disassemblers and binary
lifters~\cite{Paleari:2010:NDD:1831708.1831741,Kim:2017:TIR:3155562.3155609},
and very recently to test JavaScript
debuggers~\cite{DBLP:conf/sigsoft/LehmannP18}. All in all, it has
shown to be flexible and effective for a wide range of
applications. Surprisingly, not much work has been done on
differential testing of \js\ engines. Mozilla uses differential
testing to look for discrepancies across different configurations of
the same version of its \smonkey\ engine (using the ``compare\_jit''
flag of jsfunfuzz~\cite{jsfunfuzz}) whereas we focus on discrepancy
across engines. Patra and Pradel evaluated their language-agnostic
fuzzing strategy using differential testing. Their focuses on finding
differential bugs across multiple
browsers~\cite{patra2016learning}. As such they specialized their
fuzzer to HTML and JS (see Section~\ref{sec:testing-js-engines}). In
contrast to Patra and Pradel, we did not propose new techniques; our
contribution was empirical.

@@

%% \Fix{
%% although more research needs be done to automate
%% parts of the inspection process (\eg{}, reduction of false alarms),}

JavaScript is a very popular language. For that reason, bugs in engine
implementations often affect lots of people and organizations.
Implementing correct engines is challenging because the specification
is intentionally incomplete and evolves frequently. Finding bugs in
\js\ engines is challenging for similar reasons.

This paper reports on a study to evaluate two diversity-aware
techniques--test transplantation and cross-engine differential
testing. The first technique explores diversity of test sources; it
runs the test suite of one given engine in another engine. The second
technique explores diversity of implementations; it fuzzes existing
inputs and then compares the output produced by different engines with
a differential oracle.

We found that both techniques were very effective at finding bugs in
\js\ engines.  Overall, we reported \totalBugsReported\ bugs in this
study. Of which, \totalBugsConfirmed\ were confirmed by developers and
\totalBugsFixed\ were fixed. Although more work is necessary to reduce
cost of manual analysis, we found that our results provide strong
evidence that exploring diversity should be encouraged to find
functional bugs in JavaScript engines.

In the near future, we plan to explore techniques to prioritize the
warnings reported by these techniques and to continue involving
students in the task of diagnosing these warnings. Using learning
techniques for prioritization, similar to what Chen and
colleagues~\cite{Chen:2017:LPT:3097368.3097451} did to prioritize the
warnings reported by CSmith~\cite{Yang:2011:FUB:1993498.1993532},
seems a promising starting point for this improvement. The result
data, infrastructure, and experimental scripts we prepared to run this
study will be available upon request.

@@

%%  LocalWords:  formedness Embeddings NLP Convolutional dataset DT
%%  LocalWords:  JS JIT fuzzer jsfunfuzz jit funfuzz JavaScriptCore
%%  LocalWords:  bytecodes JScript Ecma toPrecision Kangax JSVU cccrr
%%  LocalWords:  WebKit ReferenceError TypeError drainMicrotasks sep
%%  LocalWords:  lineNumber ccrrrr Duktape JSI js runtime NN arrowed
%%  LocalWords:  embeddings perceptron toolchain fuzzers eshost cli
%%  LocalWords:  pre ArrayBuffer RangeError ToIndex getInt byteOffset
%%  LocalWords:  Clusterization SpiderMonkey DataView setInt Radamsa
%%  LocalWords:  QuickFuzz Grammarinator LangFuzz Megadeth AFL RQ crr
%%  LocalWords:  libFuzzer sanitizers subtrees conformant crrrr OME
%%  LocalWords:  ome FP TP ccr FPs TPs crrrrr ccrr escapeinside NUL
%%  LocalWords:  basicstyle boxpos numberstyle morekeywords assertEq
%%  LocalWords:  typeof valueOf ascii unary NaN eval SyntaxError Yu
%%  LocalWords:  iterable instanceof foofoofoo testcase SSL TLS SIMD
%%  LocalWords:  Lidbury OpenCL GPUs Purdom parsers Lammel Shulte PHP
%%  LocalWords:  Geno XPath combinatorial CSmith LLVM disassemblers
%%  LocalWords:  Patra Pradel namespaces Eshkevari WordPress
